---
title: "bankloan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(dplyr)
library(ggplot2)
library(tidymodels)
library(tidyverse)
library(GGally)
```

## Multiple Linear Regression on Bank Loans
This project implements a MLR model checking the relationship between loan assigned to customers against explanatory variables including annual income, monthly debt and credit score
```{r load_data}
df <- read.csv("credit_test.csv")

#get columns required
selected_cols <- df[, c("Current.Loan.Amount", "Credit.Score", "Annual.Income","Monthly.Debt","Term")]
# Rename the columns to avoid having dots
colnames(selected_cols) <- c("LoanAmount", "CreditScore", "AnnualIncome", "MonthlyDebt", "Term")
# Replace text values with 1 for "short term" and 2 for "long term" in the "Term" column
selected_cols$Term <- ifelse(selected_cols$Term == "Short Term", 1, 2)
selected_cols
```

## Task 2: Overall model - is at least one predictor useful?
One predictor is not sufficient in determining the behavior of a certain parameter used in explaining a specific phenomenon.

```{r}
data <- selected_cols 
# review any visual patterns
  data %>% 
    select(LoanAmount, Term, AnnualIncome) %>% 
    ggpairs()

  #fit the mlr model
  lm_spec <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")

  mlr_mod <- lm_spec %>% 
  fit(LoanAmount ~ Term + AnnualIncome, data = data)

  tidy(mlr_mod)
```

